---
title: "Reproducible Research - Project 1"
author: "Dinesh Kumar Subbarayalu"
date: "February 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 1

```{r Loading_Library}
library(lubridate)
library(dplyr)
library(ggplot2)
```

##### Loading Dataset:
```{r Loading_Pre-Processing}
activity <- read.csv("input/activity.csv")
activity$date <- ymd(activity$date)
```

##### Sample data:
```{r Sample}
head(activity)
```

### Analysis by Day

##### Total Steps taken each Day:

```{r steps_by_day}
tot_steps <- activity %>% group_by(date) %>% summarize(tsteps=sum(steps)) %>% na.omit()
ggplot(tot_steps, aes(x=tsteps)) + geom_histogram()
```

```{r steps_by_day_stats}
mn <- mean(tot_steps$tsteps)

md <- median(tot_steps$tsteps)
```

Mean = `r mn`
Median = `r md`

### Analysis by Interval

##### Average Steps by 5 minute interval:
```{r steps_by_interval}
avg_steps <- activity %>% group_by(interval) %>% summarize(msteps=mean(steps, na.rm = TRUE))
ggplot(avg_steps, aes(x=interval, y=msteps)) + geom_line()
```

```{r max_steps_by_interval}
max_stps <- avg_steps[avg_steps$msteps==max(avg_steps$msteps), "interval"]
```

The average steps is maximum at: `r max_stps` time interval of the day

### Imputing missing values

```{r missing_check}
missingval <- sum(is.na(activity))
```

There are totally `r missingval` rows with missing values

The missing values are imputed with mean steps in that 5 minute interval

##### Sample data after imputing:
```{r impute_missing}
fill_missing <- activity %>% group_by(interval) %>% summarize(msteps=mean(steps, na.rm = TRUE)) %>%
  right_join(activity) %>% mutate(steps = ifelse(is.na(steps), msteps, steps)) %>% select(-msteps)

head(fill_missing)
```

##### Total steps taken each Day after imputing missing values:
```{r steps_by_day_after_imputing}
tot_steps <- fill_missing %>% group_by(date) %>% summarize(tsteps=sum(steps)) %>% na.omit()

mn_imp <- mean(tot_steps$tsteps)
md_imp <- median(tot_steps$tsteps)

ggplot(tot_steps, aes(x=tsteps)) + geom_histogram()
```

Mean after imputing = `r mn_imp`
Median after imputing = `r md_imp`

The mean and median are now equal after imputation. This suggests that the data is symmetrical.

### Weekday Vs Weekend Analysis

```{r weekdayVsweekend}
activity_weekday <- fill_missing %>% mutate(day = weekdays(date)) %>% 
  mutate(day_type = ifelse(((day == "Saturday") | (day == "Sunday")), "Weekend", "Weekday")) %>%
  mutate(day_type = as.factor(day_type))

head(activity_weekday)
```

##### Average interval analysis: Weekday Vs Weekend
```{r weekdayVsweekend_plot}
avg_steps <- activity_weekday %>% group_by(interval, day_type) %>% summarize(msteps=mean(steps, na.rm = TRUE))
ggplot(avg_steps, aes(x=interval, y=msteps, color=day_type)) + 
  geom_line() + 
  facet_grid(day_type ~ .) +
  labs(title = "Mean of Steps by Interval", x = "interval", y = "steps")
```

We can clearly see that, activity is higher throughout the day during weekends than in weekdays